Manuscript Draft\\ \vspace{2cm} What is long-distance dispersal? And a taxonomy of dispersal events

### 2016-06-28 Manuscript/MS- What is LDD.tex
Edits to Discussion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS Draft Journal of Ecology Special Issue, Dispersal - 
%% Pedro - 07 April 2016
%----------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
\documentclass[a4paper, 12pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\pagestyle{plain}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                                                    
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
%\usepackage{multirow}
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\usepackage{parskip}
\linespread{1.25}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}

@@

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.

Combining spatially-explicit references to the geographic population limits and the $N^b_e$ area extent helps avoiding some imprecision in setting distance thresholds to characterize LDD events. In addition, the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events. When methods available to assign frugivore taxa to the analyzed dispersal events, as in the study case with \textit{P. mahaleb}, a classification in the four categories of events is possible. The frugivore assemblage of \textit{P. mahaleb} is composed by a diversified set of animal species spanning a wide rage of body sizes, ca. 12-14000 g in weight. We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}. If the results for \textit{P. mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events and a crucial role for in-situ population recruitment, and large-bodied species with a disproportionate contribution to LDD events.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MS Draft Journal of Ecology Special Issue, Dispersal - 
%% Pedro - 07 April 2016
%----------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Headers
\documentclass[a4paper, 12pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\pagestyle{plain}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{rotating}                                                                                    
\newcommand{\tab}{\hspace{5mm}}
\usepackage{tabularx} 
%\usepackage{multirow}
\usepackage[margin=10pt,font=small,labelfont=bf]{caption}
\usepackage[left]{lineno}
\usepackage{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy}     % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{{\tiny Jordano - What is long-distance dispersal?}}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}
\usepackage{parskip}
\linespread{1.75}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title page
\begin{document}

@@

Long-distance dispersal (LDD) is a major component of the population dynamics, genetic structure, and biogeographic history of plant species. It determines the colonization ability of new habitats and the possibilities for fragmented populations to sustain a cohesive metapopulation by immigration-emigration dynamics that rely on LDD events \citep{Nathan:2008is,Schurr2009long}. Yet our current understanding of the extent, frequency, and consequences of LDD is very limited. On one hand, theoretical models fail to predict accurately the behavior of the tail of the dispersal functions, and thus fail to predict very basic properties of LDD. On the other hand, we still have very limited documentation of actual LDD events in natural populations and we still see LDD as a sporadic, rarely far-reaching process still marked with the stamp of natural history curiosity.

Combining spatially-explicit references to the geographic population limits and the genetic neighborhood area extent ($N^b_e$) helps avoiding some imprecision in setting distance thresholds to characterize LDD events \citep{Jones:2008il}. In addition, the framework outlined in Table 1 bridges the combined demographic and genetic effects of LDD events. When methods available to assign frugivore taxa to the analyzed dispersal events, as in the study case with \textit{P. mahaleb}, a classification in the four categories of events is possible.

The frugivore assemblage of \textit{P. mahaleb} is composed by a diversified set of animal species spanning a wide size range, ca. 12-14000 g in body mass. We might expect that this extreme variation translates in an ample pattern of foraging modes, movement distances, and fruit/seed processing \citep{Jordano:2000ft}. If the results for \textit{P. mahaleb} are generalizable to other disperser assemblages, it seems that the functional roles of frugivore species in terms of contributions to LDD events are structured in two distinct groups: small-bodied frugivores, with substantial contributions to SDD events, and large-bodied species with a disproportionate contribution to LDD events. Both components of this sort of diplochorous \citep{vanderWall:2004hv} dispersal system are very frequent in fleshy-fruited plants with diversified frugivore assemblages \citep{Galetti:2013jv}. In such cases, small-bodied frugivores largely contribute the short-distance dispersal key to support \textit{in situ} recruitment and population persistence. Yet the large-bodied frugivores distinctly contribute LDD events that sustain the connectivity of metapopulation scenarios \citep{Urban:2001}. Yet, as shown in Table 1, SDD and LDD events can be more complex when we consider the contributions to gene flow via seed and the consequences in terms of structure and spatial distributions of the genetic pools.

A number of classic studies have demonstrated that the activity of large furgivores may also significantly contribute to SDD events and inefficient dispersal: i.e., territorial defence, short gut retention times relative to on-tree foraging, frequent revisititation of same trees and perches, etc., resulting in substantial SDD events \citep{Pratt:1983gf,Pratt:1984ly,Snow:1984ul,Snow:1988ve,Wheelwright:1991qf}. Yet these large-bodied frugivores are crucial for both $LDD_{loc}$ and $LDD_{ss}$, given that extensive movement patterns and extremely large foraging ranges may frequently contribute dissemination beyond distance thresholds defined with either spatial landscape or genetic references. Recent analyses of the movement ecology of large frugivores, coupled with results of their seed dispersal services emphasize that LDD are by no means exceptional, either in terms of frequency and extent \citep{Westcott:2005,Bueno:2013cg}.

@@

Pollen and seed dispersal in plants are essentially spatially-structured processes for which the outcomes of interactions with dispersal vectors is intimately linked to landscape features. Given this mechanistic link between the features of the vector and the environments where its displacement occurs \citep{Nathan:2008fx}, consideration of landscape is key to understand the consequences of LDD events. Yet these consequences hit two central aspects of plant life-histories: the demographic recruitment process \citep{Harper:1977aa}, and the genetic signatures of pollen- and seed-mediated gene flow .
